Equity conceptualizations and standards within transportation literature: a scoping review

### 2023-10-24 Manuscript/Manuscript.qmd
The outstanding sub-sections are complete (they may need additional references, but read well enough).

The third population group most commonly studied is what might be called 'composite vulnerable population groups': the intersection of several individual traits. These papers use composite vulnerability indices that captures multiple population characteristics, including low-income, unemployment, immigrant status, family household characteristics, and so on. These indices are typically generated from official government sources or author-informed census data creation, and produced using a variety of methodologies. For instance, @awuorInfluenceEnvironmentalHealth2019 use the Neighbourhood Equity Index (NEI) to investigate disparities in premature death in Toronto (Canada). The NEI is a composite index that was developed by the city to capture differences in the City's neighbourhoods by ranking them based on socio-economic characteristics (e.g., social assistance, unemployment, income) and physical environmental characteristics such as green space availability. Other works use national census indicators such as the social and housing deprivation index [e.g., @pucciInequalitiesJobrelatedAccessibility2019 or explore transport disadvantage, equity in policy implementation, or transport-related mortality burden by means of census measures (e.g., household poverty) and transport-related indicators [e.g., accessibility; @@aldredEquityNewActive2021; @iungmanImpactUrbanTransport2021; @sunPublicTransportAvailability2021; @scheurerSpatialAccessibilityPublic2017]. Similarly, Environmental Justice (EJ) indicators have been used in the US literature to identify neighbourhoods that have a higher than average proportion of low-income and non-white populations (i.e., a composite vulnerable population group'). Numerous studies have used EJ analysis to evaluate the equity impacts of transportation projects [e.g., @rowangouldIdentifyingEnvironmentalJustice2016; @parkTransitParksEnvironmental2021; @reddySafeguardingMinorityCivil2010].

@@

## "What" are the objects of transportation justice
<!-- 
NOTE: THIS SECTION DISCUSSED AT THE SAME TIME "WHAT" AND "HOW", BUT I SUGGEST TO SPLIT THE DISCUSSION IN TWO PARTS, SO THAT THE "HOW" IS ON PAR WITH OTHER ELEMENTS OF THE ANALYTICAL FRAMEWORK. IF POSSIBLE, IT WOULD BE GOOD TO FURTHER DEVELOP THE "HOW" SECTION, AS IT IS THE MOST IMPORTANT OF THE REPORT!
-->
Inequities that focus on accessibility/mobility -> opportunity standards are typically travel impedance standards, and affordability thresholds (e.g.,more than 10% of monthly income should not be spent on transport [@rivasHowAffordableTransportation2018] to work and/or other necessary destinations). Oftentimes spatial equity, disparity analysis (spatial mismatch e.g., @mulleyAreNetworkPlanning2015) are operationalized. Infrastructure standards are also suggested -> often times access to destinations is conceptualized as a "right" as such being able to enter the transportation network and physically use it the pre-requisite that accessibility measures do not capture (e.g., (dis)abilities focused).

Inequities that focus on transport-related environmental externalities -> pollution/noise standards are suggested, often using WHO guidelines or local neighbourhood averages as a comparison.

Inequities that focus on human-health  -> health standards, the reliance on population transport health literature such as physical activity guideline recommendation per week or excess mortality burden.

Inequities that focus on traffic-safety -> the use of engineering standards and best practice

Of course, some papers focus on cross-dimensional transport inequity dimensions. Measures such as urban livability, etc. a more holistic conceptualization of transport equity. This is what we think is missing from the literature, and what's needed to move forward to justice.

## "How" is fairness determined

<!-- 
NOTE: MOST IMPORTANT!
-->

@@

<!--
NOTE: IS THE TEXT BELOW A LITTLE REDUNDANT AFTER MOVING THE CALLS TO THIS SECTION? AS: Agreed.

for all stakeholders transport system decision makers in defining equitable transportation standards, particularly in collecting data that can be used to highlight and track inequalities such that standards can be set.

1. **Understand what transport inequities you are measuring**. Common measures used in transport systems are accessibility measures. What does the measure mean and does potential access relate to experienced outcomes? Next, what does an un equitable and equitable arrangement look like ? This is context specific and locality sensitive. Global north has different standards than global south.

2. **Understand what institutional context we’re operating in**. For example, what is the structure of transport planning. What Does WHO guidelines make sense to use as a standard for the community's context?

3. **Understand what institutional context transfer looks like**. A standard set in one context, may not transfer to another community in the same region.

4. **And be clear with our terms – what is equity and for whom?**. If '15 minute city' is the policy goal, what conceptualization is driving this standard? Is it sufficientarism? Egalitariansm? The standard can be interpreted a multitude of ways. The conceptualization needs to guide the standard. Being explicit about the conceptualization is important.

-->

## Call 1: The need for explicit conceptualizations and grounded standards

The rationale behind the conceptualizations of standards are often implicit within the literature: for example, @muellerSocioeconomicInequalitiesUrban2018 suggests the relative risk of mortality as related to transport-related air pollution should not be higher in deprived groups than the general population. It is not evident from the paper what conceptualization drives this focus on human-health inequities but can be implicitly inferred to be general "well being". To move equity analysis outputs towards just transport futures, explicit inclusion of the rationale for justice, the "Why?", should become more common place.

Some standards are also seemingly arbitrary. For example, @caoIndicatorsSocioSpatialTransport2017 proposes 20 ferries per day to avoid social exclusion for inter-island transport planning in the Philippines in their analysis, but admits that a standard should be politically determined. It is unclear if, for instance, 10 or 30 ferries would make a difference in a specific quality-of-life outcome or if that number is tied to funding/resource constraints. If literature is to recommend standards, researchers need standards linked to grounded outcomes. For decision-makers setting standards, measuring inequities and moving towards setting flexible guidelines for standards is the next step. As such, we believe, firmer justification of standards are lacking in the reviewed literature and are needed.

## Call 2: Need for creative methods for systems-thinking approach to inequities

On the methodological side, we believe more mixed methods are needed in transport equity research.
Conceptualizations and standards are usually discussed from purely qualitative or quantitative approaches, a missed opportunity to combine the strengths of both approaches, whether by deep diving into some particular experiences or perceptions through qualitative methods or tailoring more meaningful quantitative analysis after qualitative explorations. For instance, @xieWonCycleRoute2018 finds through interviews and go-alongs with women cyclists that the standard Cycling Level of Service (CLS) tools used by engineers to plan cycling infrastructure misses the critical gendered perspective. Further, @somenahalliAgingTransportMobility2007 surveys older adults to understand their mobility issues, revealing factors that are unseen in standard daily travel surveys.

While there is plenty of disparity analysis in the reviewed literature, frequently the standards used do not explicitly reflect practical application. For example, metrics of accessibility (usually measured with travel impedance cut-offs of between 15 to 60 minutes depending on the destination, population group and mode) are used to show descriptive differences among areas and groups but with scarce practice implications. How low of an accessibility index is too low? Conversely, if a sufficient accessibility is met, high accessibility results in higher inequalities and are inequalities of a certain level an issue? This lack of discussion makes translating results from these equity analysis into policy practice that may move conditions towards justice challenging. Creative methods and discussion of metrics of quantitative accessibility with intention of practical intervention should be paired with results.

@@

## Call 4: More direct and explicit links between standards and experienced outcomes

We believe that a more robust assessment of the implications of equity standards on life outcomes is lacking in the literature and is essential for equity analysis that translate into just practice. While the estimation of the benefits of increased mobility or accessibility, or reducing affordability burdens and transport externalities is commonplace in the literature, these estimations need to be associated with outcomes like life and neighborhood satisfaction, subjective well-being, mental and physical health, social capital, among others.

## Call 5: Towards evaluations of interventions and policies in and across communities
We believe there is a need to evaluate more equity interventions or policies. In our review, only 19 out of 165 studies assess specific transport-related policy interventions through an equity lens <!-- double check this number-->. Examples include mode-shift from driving to active school travel [@mammenSchoolTravelPlanning2014], transit fare restructures [@hickeyUsingQuantitativeMethods2010] and spatial analysis of Low Traffic Neighborhoods [@aldredEquityNewActive2021]. The assessment of interventions and tracking associated outcomes can be thought of as a key step towards transport justice. Outcomes of interventions can be compared within and between communities, and cross community comparisons can be created that may expedite the adoption of effective policies that move towards just outcomes.

## Final remarks
<!-- HERE IS AN OPPORTUNITY HERE TO CIRCLE BACK TO BEZOS SUBORBITAL FLIGHT, TO CLOSE THE REPORT WITH A REMINDER OF WHY IT IS IMPORTANT TO LEARN ABOUT JUSTICE.-->
To conclude, transportation equity analysis is important. In this report, we outline a conceptual framework for structuring equity analysis that may move us closer to transport justice, summarize findings from 165 relevant transportation equity papers from the academic literature, and critically appraised these findings with the conceptual transport justice framework.

This framework can be succinctly summarized as a classification exercise that explicitly supports a "Why" rationale. In conducting an equity analysis, the subject of justice "Who" must contextualizes the process, "What" the distribution of transport-related benefits and burdens looks like, "How" these inequities are measured and benchmarked, surrounded by the "Where" across temporally changing "When"s all should be explicitly stated. In this way, the rationale for "Why" the equity analysis is important and what root cause of/for (in)justice is motivating the analysis in the first place may be creitical to ultimately mobilizing equity analyses into transport justice.

From this framework and the critical apraisal of the literature, we outline five call to actions for researchers and decision-makers that should be adopted to move equity analysis closer to transport justice. These include:

- Explicit "Why?" rationale along with supporting "Where", "When", "Who", "What: and "How" framings for equity analysis. This means specific conceptualizations and grounded standards.
- Creative methods of equity appraisal that can be paired with systems thinking; current methods are not used to support this processes.
- Data availability as a matter of justice: who is not being counted and who should be by whom and why? 
- More direct links between equity standards and experienced conditions and outcomes. 
- We shouldn't shy away from appraise of policies. Work towards analysis that can be internationally compared.

As mentioned in the introductory paragraphs, defining _fairness_ is complex, but complexity is no excuse to shirk the task. A founding rationale for the root "Why?" of injustice is needed to inform equity analysis and set standards: why is it that Bezo (and the likes) ability to fly to space while others are firmly grounded on earth on penny-pinched public transport systems to Bezo-connected job opportunities unfair? What aspects? To whom and in what ways? The rationale for "Why" is fundamental to fairness and justice. We must be explicit, intentional and clear in our work in equity analysis such that standards can be created, equity analysis be tracked, and used to inform policy intervention that move towards justice.

>>>>

The third population group most commonly studied is what might be called 'composite vulnerable population groups': the intersection of several individual traits. These papers use composite vulnerability indices that captures multiple population characteristics, including low-income, unemployment, immigrant status, family household characteristics, and so on. These indices are typically generated from official government sources or author-informed census data creation, and produced using a variety of methodologies. For instance, @awuorInfluenceEnvironmentalHealth2019 use the Neighbourhood Equity Index (NEI) to investigate disparities in premature death in Toronto (Canada). The NEI is a composite index that was developed by the city to capture differences in the City's neighbourhoods by ranking them based on socio-economic characteristics (e.g., social assistance, unemployment, income) and physical environmental characteristics such as green space availability. Other works use national census indicators such as the social and housing deprivation index [e.g., @pucciInequalitiesJobrelatedAccessibility2019 or explore transport disadvantage, equity in policy implementation, or transport-related mortality burden by means of census measures (e.g., household poverty) and transport-related indicators [e.g., accessibility; @aldredEquityNewActive2021; @iungmanImpactUrbanTransport2021; @sunPublicTransportAvailability2021; @scheurerSpatialAccessibilityPublic2017]. Similarly, Environmental Justice (EJ) indicators have been used in the US literature to identify neighbourhoods that have a higher than average proportion of low-income and non-white populations (i.e., a composite vulnerable population group'). Numerous studies have used EJ analysis to evaluate the equity impacts of transportation projects [e.g., @rowangouldIdentifyingEnvironmentalJustice2016; @parkTransitParksEnvironmental2021; @reddySafeguardingMinorityCivil2010].

@@

## "What" are the objects of transportation justice
<!-- 
NOTE: THIS SECTION DISCUSSED AT THE SAME TIME "WHAT" AND "HOW", BUT I SUGGEST TO SPLIT THE DISCUSSION IN TWO PARTS, SO THAT THE "HOW" IS ON PAR WITH OTHER ELEMENTS OF THE ANALYTICAL FRAMEWORK. IF POSSIBLE, IT WOULD BE GOOD TO FURTHER DEVELOP THE "HOW" SECTION, AS IT IS THE MOST IMPORTANT OF THE REPORT!

Here, I've decided to talk in more depth about the equity conceptualization groups (rights, vertical equity, horizontal equity, ineq. externalities, TR exclusion, wellbeing, spatial equity, utilitarianism) across each equity dimension (access/mobility, envionrmental externality, human health, and traffic-safety).

NOTE: I think we should remove 'inequitable externality" from the conceptualizations -- fold it into horizontal and/or vertical equity. Would need to update Fig 6.
-->
Defining "What" are the things measured and the inequities that characterise their distribution is fundamental to tracking the progress towards transport justice. In the corpus reviewed, we characterize the 'thing' of inequities (the "What") in the following dimensions: 1) a focus on accessibility and mobility, 2) transport-related environmental externalities, 3) centering human-health as related to transport systems, 4) human safety related to vehicular-traffic, and 5) cross-dimensional classifications.

Depending on the characterisation of "What", the thing of inequities and the object of justice, the focus varies. The literature offers a variety of conclusions and recommendations that span across the benefits of transportation (e.g., destinations types) that are often co-mingled with population groups and the tools of mobility. For instance, in a slice of the literature centered on the distributions of accessibility/mobility, some literature centers their equity analysis on populations with disabilities (the "Who") and their limitations in accessing transportation infrastructure (their mobility tools) e.g., [@jimenez-espadaResearchProblemUniversal2021; @daamenAssessingGapPublic2008; @bharathyRevisitingClearFloor2018]; the "What?" are these inequities of accessing existing transportation infrastructure (relative to other populations) and can be extended to describe the unequal application of _Rights_ in a legal sense and/or beyond as the _right_ to participate in the production of urban space [@lefebvreDroitVille1967]. Inequities in accessing infrastructure, the "What", can be seen as supported by a rationale for justice (the "Why"). In a lot of the reviewed corpus, the rationale often stopped short in advocating for the _right_ to the city, and instead frames the rational as a _legal or envisioned legal right_ (e.g., ADA regulation [@bharathyRevisitingClearFloor2018] or the goal of "access for all" in a land-use transportation master plan [@limFacilitatingIndependentCommuting2021]).

Another slice of the objects (the "What") of justice in the corpus see their distributions assessed through conceptualizations of _horizontal_ and/or _spatial_ equity. These papers typically apply quantitative methods and see _inequity_ in the distribution as the issue at hand; they are quiet in their rationales of justice. As mentioned in the summary of findings, examples often include literature that has selected some sort of travel impedance threshold (e.g, populations who cannot access hospitals within 1 hour by car [@shenMeasuringSpatialAccessibility2020]) and discusses the distribution of the calculated accessibility index spatially across the population e.g., [@monzonEfficiencySpatialEquity2013] or population groups (the "Who") e.g., [@sharmaPublicTransitAccessibility2021]. These papers may also conceptualize traffic-related air pollution, noise pollution, urban air temperature, among other transport-related externalities. Inequities in this "What" can be theorized to be equal if similar levels are attainable by all populations (horizontal equity) or all spatial areas (spatial equity). <!-- maybe change this theorization to 'equality' not horizontal/spatial equity?--> However, discussion of both minimums and maximums (those contributing the most to inequities) is seldom explicitly discussed through the lens of justice in the reviewed corpus.

<!-- maybe change the "What" to be an object of INEQUITIES? not just justice? -->
Another branch of papers theorize the object of justice to be population _well-being_ and they typically ask _what is enough to lead a satisfactory life (as related to transportation)_. These papers are typically quantitative or mixed-methods, and typically more explicitly identify the object of (in)justice, relative to papers concerned with _spatial_ and _horizontal_ inequities. For instance, as previously discussed, papers using minimum physical activity guidelines and questionnaires that seek to understand population's activity related to active transportation infrastructure. We can assume the use of mixed/qualitative methods in a study's design as well as the use of standards linked to health outcomes (e.g., minimum travel times to get emergency treatment, premature mortality) leads to papers with a firmer identification of objects of justice.

Finally, there is another branch of research that is predominately quantitative and incorporates qualitative results or uses mixed-methods. These papers conceptualize the object of justice to be _transport related social exclusion_,  _vertical equity_, and/or _sufficientarian/capabilities_. What is characteristic of these papers is the identification of standards that are linked to often linked to outcomes like papers that conceptualize _well being_, however outcomes are often more tangible or welfare informed. As discussed, disadvantaged groups are the focus of analysis and the linkages between well-being and transport-related social exclusion [e.g., @delboscTransportProblemsThat2011; @churchillTransportPovertySubjective2019], areas more likely to experience transport poverty [@allenSizingTransportPoverty2019a], food deserts [@mckeyCrowdsourcedMappingHealthy2020], or transport-related energy poverty [@robinsonDoubleEnergyVulnerability2020; @berryInvestigatingFuelPoverty2016; @berryDistributionalEffectsCarbon2019] are central to the object of justice.

<!-- In what's described, what's missing from equity analysis in their conceptualization of the "What?" is more explicit rationale for the object of justice. A strong conceptualization of the "What" can make theorizing the "Why" rationale of justice easier.  -->

## "How" is fairness determined
<!-- 
NOTE: MOST IMPORTANT!

I need to add in more references from the corpus here... ran out of time.
-->
How we define fairness should be fundamental to equity analysis. In the reviewed literature, we summarize the philosophical perspectives and "how" (by what methods and measures) they are operationalized into standards that we group as: 1) opportunity standards, 2) infrastructure standards, 3) environmental+ standards, and 4) population standards.

Papers that put forward opportunity standards typically use methods related to travel impedance (e.g., more than 10% of monthly income should not be spent on transport [@rivasHowAffordableTransportation2018] to work and/or other necessary destinations). Alternatively, they operationalize spatial mismatch (disparity analysis) [e.g., @mulleyAreNetworkPlanning2015] and identify areas of _relative_ (regionally) inequities that should be addressed base on impedance thresholds to certain destination types. Others apply these same types of methods but tend to focus of human health, traffic-related safety, or cross-dimensional "What"s. Methods can include inequality measures, poverty measures, and optimization methods to create 'optimal' landscapes, and once again vary depending on "Who" and "What" is centered.

Infrastructure standards are also suggested and used implicitly or explicitly as standards for fairness. Often times, access to destinations is conceptualized as a _Right_ or a _right_, as such being able to enter the transportation network and physically use it the pre-requisite that accessibility measures do not capture (e.g., (dis)abilities focused). But reflected in the selection of a wide variety of approaches, the way the equity analysis is approached (who is centered and what is counted) informs _how_ fairness is assessed; be it an established engineering standard, a critic of an engineering standard, or a community-informed infrastructure need.

Papers that propose transport-related environmental externalities standards are typically informed by international or regional health- or environmental- standards. A commonly used guidance document are WHO standards, be it traffic-related pollution, or noise standards and SGS goals on urban sustainability. The advantage of the use of these international standards is that they can facilitate cross community and temporal comparison to track progress towards equity targets; though they have limitations, as they have been critiques for being too general for practical application. Overall, the ability to measure externalities tangibly (i.e., the road-side noise pollution produced by vehicles, the reduction of emissions associated with public transit modal shift) create interpretable numbers that can be measured and compared against to track progress towards these goals. These tangible metrics may be easier to interpret and compare across communities than using accessibility/mobility measures (being their methods are abundant, and contain embedded differing assumptions depending on the application).

In terms of population standards, papers that operationalize them in their analysis typically focus on health-informed standards (e.g., physical activity guidelines [@timperioSafetyNumbersDoes2015]) or identify a significant population-based threshold (e.g., an EJ community as defined by U.S. EJ analysis [@rowangouldIdentifyingEnvironmentalJustice2016]).

We believe that all forms of these standards, when linked to tangible outcomes can be used to track "fairness" in a region and can be used to track multi-dimensional progress towards justice depending on how robustly these standards are justified. Justification is exceptionally important, and we believe that a lot of the literature reviewed may miss this mark. However, literature that does not is, more often than not, connects to qualitative results and/or operationalizes mixed-methods.

@@

## Call 1: The need for explicit conceptualizations and grounded standards
<!--*And be clear with our terms – what is equity and for whom?**-->
The rationale behind the conceptualizations of standards are often implicit within the literature: for example, @muellerSocioeconomicInequalitiesUrban2018 suggests the relative risk of mortality as related to transport-related air pollution should not be higher in deprived groups than the general population. It is not evident from the paper what conceptualization drives this focus on human-health inequities but can be implicitly inferred to be general "well being". To move equity analysis outputs towards just transport futures, explicit inclusion of the rationale for justice, the "Why?", should become more common place.

Some standards are also seemingly arbitrary. For example, @caoIndicatorsSocioSpatialTransport2017 proposes 20 ferries per day to avoid social exclusion for inter-island transport planning in the Philippines in their analysis, but admits that a standard should be politically determined. It is unclear if, for instance, 10 or 30 ferries would make a difference in a specific quality-of-life outcome or if that number is tied to funding/resource constraints. As another example, if '15 minute city' is the policy goal, what conceptualization is driving this standard? Is it sufficientarism? Egalitariansm? The standard can be interpreted a multitude of ways. The framing used to conceptualize justice must guide the standard. Being explicit about justifications for the framing is important.

If literature is to recommend standards, researchers need standards linked to grounded outcomes. For decision-makers setting standards, measuring inequities and moving towards setting flexible guidelines for standards is the next step. As such, we believe, firmer justification of standards are lacking in the reviewed literature but necessary.

## Call 2: Need for creative methods for systems-thinking approach to inequities
<!--**Understand what transport inequities you are measuring**-->
On the methodological side, we believe more mixed methods are needed in transport equity research.
Conceptualizations and standards are usually discussed from purely qualitative or quantitative approaches, a missed opportunity to combine the strengths of both approaches, whether by deep diving into some particular experiences or perceptions through qualitative methods or tailoring more meaningful quantitative analysis after qualitative explorations. For instance, @xieWonCycleRoute2018 finds through interviews and go-alongs with women cyclists that the standard Cycling Level of Service (CLS) tools used by engineers to plan cycling infrastructure misses the critical gendered perspective. Further, @somenahalliAgingTransportMobility2007 surveys older adults to understand their mobility issues, revealing factors that are unseen in standard daily travel surveys.

While there is plenty of disparity analysis in the reviewed literature, frequently the standards used do not explicitly reflect practical application. For example, metrics of accessibility (usually measured with travel impedance cut-offs of between 15 to 60 minutes depending on the destination, population group and mode) are used to show descriptive differences among areas and groups but with scarce practice implications. How low of an accessibility index is too low? Conversely, if a sufficient accessibility is met, high accessibility results in higher inequalities and are inequalities of a certain level an issue? This lack of discussion makes translating results from these equity analysis into policy practice that may move conditions towards justice challenging. Creative methods and discussion of metrics of quantitative accessibility with intention of practical intervention should be paired with results; these methods should yield interpretable results and the explicit discussion of minimums _and_ maximums values in the distribution of the object of justice (the "What"), as applicable, is critical.

@@

## Call 4: More direct and explicit links between standards and experienced outcomes 
<!--*Understand what institutional context we’re operating in**-->
We believe that a more robust assessment of the implications of equity standards on life outcomes is lacking in the literature and is essential for equity analysis that translate into just practice. While the estimation of the benefits of increased mobility or accessibility, or reducing affordability burdens and transport externalities is commonplace in the literature, these estimations need to be associated with outcomes like life and neighborhood satisfaction, subjective well-being, mental and physical health, social capital, among others. Further, the use of standards that are tangibly linked to outcomes can more concretely define the objects of justice (the "What"). They may also be used to track progress towards goals across time and space.

## Call 5: Towards evaluations of interventions and policies in and across communities
<!--**Understand what institutional context transfer looks like**-->
We believe there is a need to evaluate more equity interventions or policies. In our review, only 19 out of 165 studies assess specific transport-related policy interventions through an equity lens <!-- double check this number-->. Examples include mode-shift from driving to active school travel [@mammenSchoolTravelPlanning2014], transit fare restructures [@hickeyUsingQuantitativeMethods2010] and spatial analysis of Low Traffic Neighborhoods [@aldredEquityNewActive2021]. The assessment of interventions and tracking associated outcomes can be thought of as a key step towards transport justice. Outcomes of interventions can be compared within and between communities, and cross community comparisons can be created that may expedite the adoption of effective policies that move towards just outcomes.

## Final remarks
<!-- HERE IS AN OPPORTUNITY HERE TO CIRCLE BACK TO BEZOS SUBORBITAL FLIGHT, TO CLOSE THE REPORT WITH A REMINDER OF WHY IT IS IMPORTANT TO LEARN ABOUT JUSTICE.

And this is my attempt - more time is needed but this first attempt is okay! -->
To conclude, the analysis of transportation equity is vitally important for planning: as transportation systems and technologies converge space-time, they by definition leave certain populations, depending on the area and time scale, more or less burdened or benefited. The analysis and tracking of equity implications is fundamental in formulating standards that are flexible to the fluid demands of transport justice. In this spirit, this report outlines a conceptual framework for structuring equity analysis that may move us closer to transport justice, summarize findings from 165 relevant transportation equity papers from the academic literature, and critically appraised these findings with the conceptual transport justice framework.

The transportation justice framework can be seen as a flexible classification exercise that is suppored by a "Why" rationale. In conducting an equity analysis, the subject of justice (the "Who?") must contextualize the process, the distribution of the object of justice (the "what?") must be measured, be it transport-related benefits, burdens or both, "How?" these inequities are measured and benchmarked must be described, and these questions must be contextualized by the "Where?" across temporally changing "When?"s. The explicit acknowledge and answers to these questions clarify the boundaries of the equity analysis. In this way, the rationale for "Why?" the equity analysis is important and what root cause of/for (in)justice is motivating the analysis in the first place may be critical to ultimately mobilizing equity analyses into experienced transport justice.

From this framework and the critical appraisal of the literature, we outline five call to actions for researchers and decision-makers that should be adopted to move equity analysis closer to transport justice. These include:

- Explicit "Why?" rationale along with supporting "Where", "When", "Who", "What: and "How" framings for equity analysis. This means specific conceptualizations and grounded standards.
- Creative methods of equity appraisal that can be paired with systems thinking; current methods are not used to support this processes.
- Data availability as a matter of justice: who is not being counted and who should be by whom and why? 
- More explicit links between equity standards and experienced conditions and outcomes. 
- Encouragement to continue work towards analysis and comparison of applied interventions, across communities and internationally.

As mentioned in the introductory paragraphs, defining _fairness_ is complex, but complexity is no excuse to shirk away from the task. A founding rationale for the root "Why?" of in/justice is needed to inform equity analysis and set standards: why is it that Bezo (and the likes) ability to fly to space compared to the majority of others who are firmly grounded on earth on penny-pinched public transport systems traveling to Bezo-connected job opportunities felt as _unfair_? What aspects are _unfair_? To whom and in what ways? The answer to "Why?" is fundamental in pinning down fairness and justice. We must be explicit, intentional and clear in our work in transport equity analysis such that standards can be created, equity analysis be tracked and used to inform policy intervention that move towards justice.
